---
title: "Hotel Group analysis"
author: "David"
date: "28/11/2021"
output: html_document
---

```{r libraries setup, echo=FALSE, message=FALSE, warning=FALSE}
if (!require('ggplot2')) install.packages('ggplot2'); library('ggplot2')
if (!require('knitr')) install.packages('knitr'); library('knitr')
if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
if (!require('tidyr')) install.packages('tidyr'); library('tidyr')
if (!require('here')) install.packages('tidyr'); library('here')
if (!require('readr')) install.packages('readr'); library('readr')
hotel_df <- read.csv("https://d3c33hcgiwev3.cloudfront.net/GL0bk8O2Sja9G5PDtko2uQ_31e445d7ca64417eb45aeaa08ec90bf1_hotel_bookings.csv?Expires=1638230400&Signature=USt5k8o~ozCdZsiMpeoLs36t0FSDYc5Ie8eDBabhXGMe7XBSUAVl5f4OLYF1OREYGNBj6OnFVpwO62yzrl6OuV2IsuWg9bimdg~kZzotnJ1q2YbuE-z64SJvic1q2i0sZiTtCb-KLQXATDczdDCIwqVf3l36tI0rGyQzocTNgS4_&Key-Pair-Id=APKAJLTNE6QMUY6HBC5A")
```

## Introduction:

In this notebook are reports and observations from a worldwide group of hotels.

In detail I will report:

1. The types of hotels the group has
2. The busiest months of the year for each type of hotel
3. The Ranking of Hotel Packages sold
4. Which market segment makes the most reservations per hotel type

<br>

## Let's start 

### 1. The types of hotels the group has:

```{r types of hotels, echo=FALSE, message=FALSE, warning=FALSE}
hotel_types <- unique(hotel_df$hotel)
hotel_types_df <- data.frame(hotel_types)
colnames(hotel_types_df) <- c("Types of Hotels")
kable(hotel_types_df) 
```
<br>

### 2. The busiest months of the year for each type of hotel

```{r busiest periods for city hotels, echo=FALSE, message=FALSE, warning=FALSE}
city_hotels <- hotel_df %>% filter(hotel == "City Hotel")
ggplot(city_hotels) + geom_bar(mapping=aes(x=arrival_date_month), fill="dodgerblue1") + 
  theme(axis.text.x = element_text(angle = 45)) + ggtitle("City Hotels") + xlab("") + 
  ylab("Number of guest arrivals")
city_hotels_months_count <- city_hotels %>% count(arrival_date_month)
city_hotels_months_count_df <- data.frame(city_hotels_months_count)
colnames(city_hotels_months_count_df) <- c("Months", "Arrivals")
kable(city_hotels_months_count_df %>% arrange(-Arrivals)) 
```
```{r busiest periods for resort hotels, echo=FALSE, message=FALSE, warning=FALSE}
resort_hotels <- hotel_df %>% filter(hotel == "Resort Hotel")
ggplot(resort_hotels) + geom_bar(mapping=aes(x=arrival_date_month), fill="springgreen4") + 
  theme(axis.text.x = element_text(angle = 45)) + ggtitle("Resort Hotels") + xlab("") + 
  ylab("Number of guest arrivals")
resort_hotels_months_count <- resort_hotels %>% count(arrival_date_month)
resort_hotels_months_count_df <- data.frame(resort_hotels_months_count)
colnames(resort_hotels_months_count_df) <- c("Months", "Arrivals")
kable(resort_hotels_months_count_df %>% arrange(-Arrivals))
```
<br>

### 3. The Ranking of Hotel Packages sold

```{r top 3 food packages, echo=FALSE, message=FALSE, warning=FALSE}
# unique(hotel_df$meal)
ranking_of_packages <- c("Nº 1", "Nº 2", "Nº 3", "Nº 4", "Nº 5")
packages <- c("BB", "HB", "SC", "Undefined", "FB")
packages_meaning <- c("Bed and Breakfast", "Half Board", "Self Catering", "Undefined", "Full Board")
packages_meaning_in_english <- c("Bed, Breakfast", "Bed, Breakfast, Dinner", "Bed", "Undefined", "Bed, Breakfast, Lunch, Dinner")
packages_units_sold <- c(92310, 14463, 10650, 1169, 798)
meal_plans_df <- data.frame(ranking_of_packages, packages, packages_meaning, packages_meaning_in_english, packages_units_sold)
colnames(meal_plans_df) <- c("Ranking", "Hotel packages", "Meaning of Hotel packages", "What's included on each package", "Units sold in 2015, 2016 and 2017")
kable(meal_plans_df)
```
<br>

### 4. Which market segment makes the most reservations per hotel type

```{r top market segment, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(hotel_df) + geom_bar(mapping=aes(x= market_segment, fill= market_segment)) + 
  theme(axis.text.x = element_text(angle=90)) + labs(fill= "Market Segment", x= "", y="Number of reservations") + 
  facet_wrap(~hotel)
```

<br>

## Observations:

* From the month of May until end of August, both types of hotels will receive more reservations.
* From the month of November until end of January, both hotel types will receive less reservations.
* The package most sold with a difference of 538 % relative to the second place is "Bed and breakfast"(BB).
* The market segment which provides the most reservations is Online TA (online travel agency).

<br>

## Suggested actions to the managment team:

* Recruit more staff in the whole company, during the months of February and March, to train them in order to tackle the busy periods of May, June, July and August.
* Raise the prices in the whole company, by 32% during the months May and June, and raise by 57% during the  months of July and August.
* Liaise with the marketing team in order to create strategies for the whole company that promote higher number of reservations during the months November, December and January.
* Liaise with the marketing and F&B teams to create promotions to develop the Lunch time in order to retain more reservations.
* Liaise with the I.T and marketing teams to develop the company website in order to attract more direct reservations through the company's website.